# Methods

```{r setup-methods, echo = FALSE, message = FALSE, warning = FALSE}
library(pta)
library(knitr)

opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

specdb <- src_sqlite('extdata/leaf_spectra.db')
```

## Data

Data for this analysis comes from several different projects, reported in Table \ref(projecttable).

```{r projecttable}

data(spectra_info)

byproject <- spectra_info %>%
    group_by(projectcode, projectdescription) %>%
    summarize(sites = n_distinct(sitecode),
              species = n_distinct(speciescode),
              samples = n_distinct(samplecode)) %>%
    arrange(desc(samples))

totals <- spectra_info %>%
    summarize(sites = n_distinct(sitecode),
              species = n_distinct(speciescode),
              samples = n_distinct(samplecode)) %>%
    mutate(projectcode = '', projectdescription = 'Total')

projecttable <- bind_rows(byproject, totals) %>%
    rename(code = projectcode,
           Description = projectdescription,
           '# sites' = sites, 
           '# species' = species,
           '# spectra' = samples)

kable(projecttable, caption = 'Project information')

```

```{r sitemap, fig.cap = sitemap_cap, fig.width = 10, fig.height = 7}

sitemap_cap <- paste('Sample distribution in geographic and climate space')

data(site_info)

sitemap_dat <- spectra_info %>%
    left_join(site_info) %>%
    count(sitecode, site_lat, site_lon, temp, prec, projectcode) %>%
    ungroup() %>%
    group_by(projectcode) %>%
    mutate(nproj = sum(n)) %>%
    ungroup() %>%
    mutate(projectcode = forcats::fct_reorder(projectcode, nproj)) %>% 
    arrange(desc(nproj))

plt <- ggplot(sitemap_dat) + 
    aes(x = site_lon, y = site_lat, fill = projectcode, size = nproj) +
    borders('world') +
    geom_point(pch = 21, color = 'black') + 
    scale_fill_brewer(palette = 'Paired') + 
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank()) + 
    guides(size = FALSE,
           fill = guide_legend(title = 'Project', 
                               override.aes = list(size = 4)))
plot(plt)
ggsave(filename = 'figures/sitemap.png', 
       plot = plt, 
       device = 'png',
       scale = 0.5)

plt <- ggplot() + 
    aes(x = prec, y = temp) +
    xlab('Mean annual precipitation (mm)') +
    ylab(expression(paste('Mean annual temperature (', degree, 'C)'))) +
    geom_polygon(data = biomes, aes(fill = biome), 
                 size = 0.5, alpha = 0.2, color = 'black') + 
    geom_point(data = sitemap_dat, aes(size = n, color = projectcode)) + 
    geom_label(data = biome_labs, aes(x = x, y = y, label = biome)) + 
    scale_fill_manual(values = biome_colors) + 
    scale_color_brewer(palette = 'Paired') +
    theme_bw() +
    guides(fill = FALSE,
           color = guide_legend(title = 'Project', 
                                override.aes = list(size = 4)))
plot(plt)
ggsave(filename = 'figures/biomes.png',
       plot = plt,
       device = 'png',
       scale = 0.5)

```

## Trait estimation from spectra

### PROSPECT 5B model

The PROSPECT 5B model predicts leaf reflectance and transmittance as a function of the effective number of leaf mesophyll layers, chlorophyll and carotenoid concentrations, **Cbrown**, water content, and dry matter content [@feret_prospect-4_2008], collectively referred to henceforth as "leaf optical traits".

Leaf mesophyll number is important because...

Chlorophyll concentrations control light absorption and photosynthetic potential (**Richardson et al. 2002**), and, as the most nitrogen-rich component of leaves, provide a useful indication of leaf nitrogen (and general nutrition) status [@moran_differentiation_2000].

Carotenoid pigments in this context encompass many different pigments related to the xanthophyll cycle. They are related to photochemical stress 
[@magney_assessing_2014] (** See references in Magney**) **and other stuff**.
Brown matter concentrations are **???**.

Leaf water content is related to drought stress [@carter_responses_1993; @asner_progressive_2016; @penuelas_reflectance_1993; @penuelas_reflectance_1994] (**Refs in Asner 2016**), nutrient stress (**REFS**), and insect damage [@cheng_continuous_2010] and is a key variable for assessing susceptibility of ecosystems to fire [@agee_foliar_2002].

Leaf dry matter content:

* Correlated with specific leaf area and leaf nitrogen [@laughlin_multi-trait_2010; @pierce_allocating_2013], and therefore an important indicator of plant position on the leaf economic spectrum [@wright_worldwide_2004; **others**].

In addition to exploring these optical traits, we also examine their relationship with to leaf nitrogen per unit mass, related to carboxylation efficiency and which are an important indicator of nutrient stress (**Field & Mooney 1986**), and leaf mass per area (LMA), an indicator of plant strategy with regard to carbon allocation and light availability [@ackerly_functional_2004; @poorter_causes_2009].


### Inversion algorithm

Bayesian inversion of PROSPECT [@shiklomanov_quantifying_2016].
Added effective sample size normalization, which down-weights the likelihood and therefore provides more conservative uncertainty estimates.


## Statistical analysis


Ideas:

* Calculate some 'T-statistics', as in @violle_return_2012 ?

* Try the "fourth corner method" as in @dray_testing_2008

    - Build a matrix of species abundance at different sites, a matrix of trait values of species, and a matrix describing the environmental conditions at each site

    - It seems like this falls into the Jim Clark aggregation trap of discarding individual variability. 
      Could itself be an interesting finding?
      Probably too much for one paper.

