# Results

## RTM inversion / Methodological considerations

```{r densfig, fig.cap = sumstatfigcap}
sumstatfigcap <- paste('Histograms of inversion results by model and parameter')
rsub <- results %>% 
    filter(parameter %in% params) %>%
    select(samplecode, modelname, parameter, parametermean) %>%
    collect(n = Inf) %>%
    mutate(parameter = factor(parameter, params))
gpsub <- ggplot(rsub) + aes(x = parametermean) + 
    facet_grid(modelname ~ parameter, scales = 'free_x') + 
    geom_histogram()
gpsub
rsub %>%
    group_by(modelname, parameter) %>%
    summarize(stat = sprintf('%s (%s)',
                             signif(mean(parametermean, na.rm = TRUE), 3),
                             signif(sd(parametermean, na.rm = TRUE), 3))) %>%
    spread(parameter, stat) %>%
    kable(caption = 'Summary statistics by parameter and model')
```

NOTE: Not all inversions have finished running yet, and different models are in different states of completeness, so these results should be taken with a grain of salt.

In general, none of the distributions of the means look extreme, and for the most part, they all have the same general shape. 


```{r validation, fig.cap = valid_cap}

valid_cap <- 'Comparison of spectral trait estimates with direct measurements'

convert2si <- function(value, parameter) {
    case_when(parameter %in% c('Cab', 'Car') ~ 
                udunits2::ud.convert(value, 'ug cm-2', 'kg m-2'),
              parameter %in% c('Cw', 'Cm') ~ 
                udunits2::ud.convert(value, 'g cm-2', 'kg m-2'),
              TRUE ~ NA_real_)
}

valid_plot <- valid_dat %>%
    filter((parameter == 'Cab' & trait == 'leaf_chltot_per_area') |
           (parameter == 'Car' & trait == 'leaf_cartot_per_area') |
           (parameter == 'Cw' & trait == 'leaf_water_thickness') |
           (parameter == 'Cm' & trait == 'leaf_mass_per_area'),
       modelname == 'PROSPECT 5',
       !(projectcode %in% c('foster_beetle', 'wu_brazil', 'nasa_hyspiri'))) %>%
    mutate_at(vars(matches('parameter[[:alpha:]]+')),
              convert2si, parameter = .$parameter) %>%
    mutate(month = lubridate::month(collectiondate),
           growing_season = between(month, 5, 9)) %>%
    filter(!(parameter == 'Cab' & parametermean > 0.00125),
           !(parameter == 'Cm' & parametermean > 0.25),
           !(trait == 'leaf_water_thickness' & traitvalue > 1)) %>%
    ggplot(aes(x = parametermean, 
               xmin = parametermean - parametersd,
               xmax = parametermean + parametersd,
               y = traitvalue)) + 
        facet_wrap(~parameter + trait, scales = 'free') +
        geom_errorbarh(color = 'grey', alpha = 0.25, size = 0.5) +
        geom_point(aes(color = projectcode), size = 0.8, alpha = 0.5, stroke = 0) + 
        scale_fill_brewer(palette = 'Dark2') +
        geom_abline(linetype = 'dashed') + 
        theme_bw() +
        guides(color = guide_legend(override.aes = list(size = 4, 
                                                        alpha = 1, 
                                                        pch = 15))) +
        labs(x = 'PROSPECT 5B estimate',
             y = 'Measured value')

traits <- tbl(specdb, 'trait_data') %>%
    collect(n = Inf) %>%
    mutate(trait = recode(trait, `leaf_mass_per_per_area` = 'leaf_mass_per_area')) %>%
    group_by(samplecode, trait) %>%
    summarize(traitvalue = mean(traitvalue)) %>%
    ungroup() %>%
    right_join(tbl(specdb, 'samples') %>% collect(n = Inf)) %>%
    spread(trait, traitvalue) %>%
    inner_join(modagree_dat) %>%
    mutate_at(vars(matches('Ca(b|r)')),
              funs('SI' = udunits2::ud.convert(., 'ug cm-2', 'kg m-2'))) %>%
    mutate_at(vars(matches('C(w|m)')),
              funs('SI' = udunits2::ud.convert(., 'g cm-2', 'kg m-2')))

valid_plot <- function(measure, estimate, modcodes, data, ...) {
    ny <- length(modcodes)
    oldpar <- par(mfrow = c(1, ny), mar = c(3, 3, 2, 1), oma = c(2, 2, 2, 0))
    for (i in seq_along(modcodes)) {
        estmu <- paste('mu', estimate, modcodes[i], 'SI', sep = '_')
        esthi <- paste('hi', estimate, modcodes[i], 'SI', sep = '_')
        estlo <- paste('lo', estimate, modcodes[i], 'SI', sep = '_')
        form <- formula(paste(measure, estmu, sep = ' ~ '))
        fit <- lm(form, data = data)
        r2 <- summary(fit)$r.squared
        plot(form, data = data, main = names(modcodes)[i], type = 'n')
        segments(y0 = data[[estmu]], x0 = data[[estlo]], x1 = data[[esthi]], col = 'grey80')
        points(form, data = data, pch = 19, col = 'grey50')
        abline(0, 1, col = 'red', lty = 'dashed', lwd = 2)
        abline(fit, col = 'blue', lty = 'solid', lwd = 2)
        legend('bottomright', paste('R2 = ', signif(r2, 3)), bty = 'n')
    }
    mtext('PROSPECT inversion estimate', side = 1, outer = TRUE)
    mtext('Observed', side = 2, outer = TRUE)
    mtext(paste(estimate, 'validated against', measure), side = 3, outer = TRUE)
}
    
valid_plot('leaf_chltot_per_area', 'Cab', modcodes, traits)
valid_plot('leaf_cartot_per_area', 'Car', modcodes[-1], traits)
valid_plot('leaf_water_thickness', 'Cw', modcodes, 
           filter(traits, leaf_water_thickness < 1))
valid_plot('leaf_mass_per_area', 'Cm', modcodes, traits)
```

Overall, PROSPECT 5B had the best inversion accuracy, with the highest R2 values for all parameters except dry matter content ("Cm").
By parameter, the best inversion accuracy was achieved for leaf water content ("Cw"), with all remaining parameters having roughly comparable, poor inversion accuracy.

```{r validationggplot, fig.cap = valid_ggplot_cap}

valid_ggplot_cap <- paste('Comparison of spectral trait estimates with direct', 
                          'measurements, by project and species attributes')

mod <- 'p5b'

gg_valid <- tbl(specdb, 'species') %>%
    inner_join(tbl(specdb, 'species_attributes')) %>%
    collect(n = Inf) %>%
    right_join(traits) %>%
    mutate(lg = paste(leaf_type, growth_form))

valid_vec <- c('Cab' = 'leaf_chltot_per_area',
               'Car' = 'leaf_cartot_per_area',
               'Cw' = 'leaf_water_thickness',
               'Cm' = 'leaf_mass_per_area')

valid_ggplot <- function(measure, estimate, data, ...) {
    estmu <- paste('mu', estimate, mod, 'SI', sep = '_')
    estlo <- paste('lo', estimate, mod, 'SI', sep = '_')
    esthi <- paste('hi', estimate, mod, 'SI', sep = '_')
    indat <- filter_(data, .dots = sprintf('!is.na(%s)', 
                                           c(measure, estmu, estlo, esthi)))
    xrange <- c(0, max(indat[[estmu]]))
    yrange <- c(0, max(indat[[measure]]))
    ggplot(indat) + 
        aes_string(x = estmu, y = measure, xmin = estlo, xmax = esthi, ...) + 
        geom_errorbarh(alpha = 0.15, size = 0.5) +
        geom_point(alpha = 0.5, size = 1, shape = 21, color = 'black') + 
        geom_smooth(method = 'lm', se = FALSE) + 
        geom_abline(intercept = 0, slope = 1, linetype = 'dashed') + 
        scale_color_brewer(palette = 'Dark2') + 
        scale_fill_brewer(palette = 'Dark2') +
        coord_cartesian(xlim = xrange, ylim = yrange)
}

for (i in seq_along(valid_vec)) {
    v <- valid_vec[i]
    nv <- names(valid_vec)[i]
    dat <- gg_valid
    if (v == 'leaf_water_thickness') {
        dat <- filter(dat, leaf_water_thickness < 1)
    }
    plot(valid_ggplot(v, nv, dat, color = 'projectcode', fill = 'projectcode'))
    plot(valid_ggplot(v, nv, dat, color = 'leaf_type', fill = 'leaf_type'))
    plot(valid_ggplot(v, nv, dat, color = 'growth_form', fill = 'growth_form'))
    plot(valid_ggplot(v, nv, dat, color = 'lg', fill = 'lg'))
}

```

```{r variancecompare, fig.cap=vccap}

vccap <- 'Comparison of variance between spectral estimates and direct measurements.'

variance_compare <- function(measure, estimate, data, ...) {
    mu <- paste('mu', estimate, mod, 'SI', sep = '_')
    indat <- filter_(data, .dots = sprintf('!is.na(%s)', c(measure, mu))) %>%
        summarize_at(c(measure, mu), var) %>%
        rename_(.dots = c('measure' = measure, 'estimate' = mu))
    return(indat)
}

variance_dat <- mapply(variance_compare,
                       valid_vec,
                       names(valid_vec),
                       MoreArgs = list(data = gg_valid), SIMPLIFY = FALSE) %>% 
    bind_rows(.id = 'param') %>%
    gather(variable, variance, -param)

ggplot(variance_dat) + 
    aes(x = variable, y = variance) + 
    geom_bar(stat = 'identity') + 
    facet_wrap(~param, scales = 'free')
```

In all cases, the measured values had a higher variance than estimates from spectral inversion. 
The largest difference in variance between measured values and spectral estimates was for carotenoid and dry matter contents, while variance was closest for chlorophyll content.
It should be noted that this comparison only applies to spectral values which have associated trait measurements -- i.e. the actual variance for all traits estimated from spectra may be considerably different.

```{r covariancecompare, fig.cap = cvccap}

cvccap <- paste('Comparison of covariance between spectral estimates ', 
                'and direct measurements')
cvparcols <- paste('mu', names(valid_vec), mod, 'SI', sep = '_')
names(cvparcols) <- names(valid_vec)
cvdat <- gg_valid %>%
    filter_(.dots = paste(sprintf('!is.na(%s)', valid_vec), collapse = ' | '))

mycorplot <- function(data, title, cols, ...) {
    mat <- cor(select_(data, .dots = cols), use = 'pairwise.complete.obs')
    cols <- colorRampPalette(c('red', 'grey30', 'blue'))(20)
    corrplot::corrplot.mixed(mat, lower = 'ellipse', upper = 'number', 
                             main = title, mar = c(0, 0, 1, 0), 
                             col = cols, ...)
}
par(mfrow = c(1, 2))
mycorplot(cvdat, 'Measured', valid_vec)
mycorplot(cvdat, 'Estimate', cvparcols)

```

In all cases, covariance between traits is comparable or stronger in the measured values than in the spectral inversion estimates.
Note that these results may not be completely reliable because the spectral estimates have no missing values (though they are filtered to the samples that have at least one trait measurement) whereas the trait data do.

## Sources of trait variability 

_NOTE: All of the results in this section assume that the values above are accurate.
They may change if I revise the model used to estimate spectra._

### Correlations among PROSPECT traits

```{r pcaall, fig.cap = pcaall_cap}

pcaall_cap <- paste('Principal component analysis on all PROSPECT traits')

pca_pars <- paste('mu', params, mod, sep = '_')
names(pca_pars) <- params

pcaall_dat <- gg_valid %>%
    rename_(.dots = pca_pars) %>%
    #mutate_at(params, log) %>%
    filter_(sprintf('!is.na(%s)', params))

draw_pca <- function(dat, char) {
    fit <- princomp(dat %>% select_(.dots = params), cor = TRUE)
    biplot(fit, xlabs = rep(char, nrow(dat)))
}

draw_pca(pcaall_dat, '.')

```

```{r pcasp, fig.cap = pcasp_cap}

pcasp_cap <- paste('Principal components analysis on species means.')
pcasp_dat <- pcaall_dat %>%
    group_by(speciescode) %>%
    summarize_at(params, mean)
draw_pca(pcasp_dat, 'x')

```

```{r pcares, fig.cap = pcares_cap}

pcares_cap <- paste('Principal components analysis on residuals',
                    'from species means')
pcares_dat <- pcaall_dat %>%
    group_by(speciescode) %>%
    mutate_at(params, (function(.) . - mean(.))) %>%
    ungroup
draw_pca(pcares_dat, '.')

```

The within- and across-species relationships between parameters are different.
Across species, there are two orthagonal axes of variability, with the primary axis driven by pigment concentrations and the secondary axis driven by leaf structure (as manifested by contents of brown and dry matter contents, water thickness, and number of mesophyll layers).
Within species, the primary axis of variability is **???**

### Drivers of variability in PROSPECT traits

```{r anovaplot, fig.cap = anovaplot_cap, fig.width = 10, fig.height = 10}

anovaplot_cap <- paste('Partitioning variability in PROSPECT traits')

anovadat <- pcaall_dat %>% 
    left_join(tbl(specdb, 'plots') %>% 
              select(sitecode, plotcode) %>% 
              collect) %>% 
    left_join(site_info)

vars <- tribble(
    ~variable, ~color, 
    'leaf_type', 'green4',
    'phenology', 'brown',
    'growth_form', 'purple',
    'myco_asso', 'orange',
    'shade_tolerance', 'yellow',
    'speciescode', 'green1',
    'sitecode', 'red',
    'plotcode', 'cyan')
          
anova_param <- function(dat, param, vars) {
    form <- formula(paste(param, '~', paste(vars, collapse = ' + ')))
    fit <- lm(formula = form, data = dat)
    anova_raw <- anova(fit)
    anova_dat <- anova_raw[, 'Sum Sq', drop = FALSE]
    colnames(anova_dat) <- param
    out <- as_data_frame(anova_dat) %>% rownames_to_column('variable')
    return(list(table = out, fit = fit))
}

anova_list <- lapply(params, function(x) anova_param(anovadat, x, vars$variable))
names(anova_list) <- params

anova_dat <- Reduce((function(...) merge(..., by = 'variable')), 
                    lapply(anova_list, '[[', 'table')) %>%
    gather(parameter, value, -variable) %>%
    mutate(variable = factor(variable, rev(c(vars$variable, 'Residuals'))),
           parameter = factor(parameter, params)) %>%
    group_by(parameter) %>%
    mutate(norm_value = value / sum(value))

ggplot(anova_dat) + 
    aes(x = parameter, y = norm_value, fill = variable) + 
    geom_bar(stat = 'identity') + 
    scale_fill_manual(values = rev(c(vars$color, 'grey')))
    
```

Related to the results above, the primary drivers of variability in PROSPECT traits are trait-dependent.

The number of leaf mesophyll layers, brown matter content, and water thickness are dominated largely by among-species variability.
Much of the among-species variability in the N parameter is explained by leaf type (broad vs. needle), with some additional contribution of shade tolerance.
Leaf type is also important in explaining variability in leaf water content, as is plant growth form.
However, large fractions of this variability are unexplained by any functional attribute, contributing to high residual species-level variability.

Most of the variability in the remaining traits -- pigment and dry matter contents -- is within species, and the majority of that variability remains unexplained.

For all traits, site is an important factor, accounting for approximately 25% of the variance on average.

```{r pftcoefs, fig.cap = pftcoefs_cap, fig.width = 12, fig.height = 12}

coeftab <- function(anova_list, param, variable) {
    fit <- anova_list[[param]][['fit']]
    sfit <- summary(fit)
    fitnames <- names(fit$coefficients)
    coef_inds <- grep(variable, fitnames)
    coef_mean <- sfit$coefficients[coef_inds, 'Estimate']
    coef_se <- sfit$coefficients[coef_inds, 'Std. Error'] 
    coeftab <- tibble(parameter = param,
                      variable = variable,
                      varvalue = gsub(variable, '', fitnames[coef_inds]),
                      coef_mean = coef_mean,
                      coef_se = coef_se)
    return(coeftab)
}

coef_list <- list()
coef_vars <- c('leaf_type', 'phenology', 'growth_form',
               'myco_asso', 'shade_tolerance')
i <- 1
for (v in coef_vars) {
    for (p in params) {
        coef_list[[i]] <- coeftab(anova_list, p, v)
        i <- i + 1
    }
}
coeftable <- bind_rows(coef_list) %>%
    mutate(parameter = factor(parameter, params))

ggplot(coeftable) + 
    aes(x = varvalue, y = coef_mean, 
        ymin = coef_mean - coef_se, ymax = coef_mean + coef_se) + 
    geom_pointrange() + 
    geom_hline(yintercept = 0, linetype = 'dashed', color = 'red') +
    facet_grid(parameter ~ variable, scales = 'free')

pftcoefs_cap <- paste('Linear model coefficients, +/- 1 standard error.', 
                      'Baselines are as follows:', 
                      'growth_form = graminoid,',
                      'leaf_type = broad,',
                      'myco_asso = AM,',
                      'phenology = deciduous,',
                      'shade_tolerance = intermediate')

```

Woody species (trees and shrubs) had higher N and lower Cm values than herbaceous species (grasses (zero) and herbs).
Needleleaf species had higher N and lower pigment, brown, and dry matter contents than broadleaf species.
Compared to species with arbuscular (AM, (zero)) mycorrhizal associations, species with ectomycorrhizal associations (EM/ECM) had significantly lower N and higher pigment, brown, and dry matter contents, while the reverse was true of species with erocoid (ERM) mycorrhizal associations.
Evergreen species had higher pigment, brown matter, and dry matter contents and lower N than deciduous species.
Strongly shade tolerant species had significantly higher N and lower pigment, brown matter, and dry matter contents than species with lower shade tolerance.
 
```{r traitsclimate, fig.cap = traitsclimate_cap, fig.width = 12, fig.height = 12}

traitsclimate_cap <- paste('Relationship of traits with climate')

traitcli_dat <- anovadat %>%
    select_(.dots = c('samplecode', 'speciescode', 
                      params, 'temp', 'prec')) %>%
    filter(!is.na(temp), !is.na(prec)) %>%
    gather(parameter, value, -samplecode, -temp, -prec, -speciescode) %>%
    gather(climvar, climval, -samplecode, -speciescode, -parameter, -value)

traitcli_plt <- ggplot(traitcli_dat) + 
    aes(x = climval, y = value) + 
    geom_point(color = 'grey50') + 
    geom_smooth(method = 'lm', se = FALSE) +
    facet_wrap(~parameter + climvar, scales = 'free')

plot(traitcli_plt)

```

```{r traitsclimatesp, fig.cap = traitsclimatesp_cap, fig.width = 12, fig.height = 12}

traitsclimatesp_cap <- paste('Relationship of traits with climate: ', 
                             'Species means.',
                             'Blue line is a linear fit,',
                             'red line is a quadratic fit.')
traitcli_plt %+% 
    (traitcli_dat %>% 
     group_by(speciescode, parameter, climvar, climval) %>%
     summarize(value = mean(value))) %+%
    stat_smooth(method = 'lm',
                formula = y ~ x + I(x^2),
                color = 'red', se = F)

```

```{r traitsclimateres, fig.cap = traitsclimateres_cap, fig.width = 12, fig.height = 12}

traitsclimateres_cap <- paste('Relationship of traits with climate:', 
                              'Normalized residuals from the species mean')
traitcli_plt %+% 
    (traitcli_dat %>% 
     group_by(speciescode, parameter, climvar) %>%
     mutate(value_mean = mean(value, na.rm = TRUE),
            value = (value - value_mean)/value_mean)) + 
    coord_cartesian(ylim = c(-1, 1))

```

Relationships between plant traits and climate occur primarily at the species level, with relatively flat relationships between climate and species mean residuals.
Leaf water thickness and carotenoid, brown matter, and dry matter contents all have a negative relationship with precipitation, while the N parameter and chlorophyll contents have a positive relationship.
All traits have weak positive relationships with temperature, except brown matter content, for which this relationship is strongly negative.

